---
title: "Take Home Exercise 03"
author: Christiana Lim
date: "10 Feb 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  warning: false
format: 
  html: 
    page-layout: full
---

# Dataset and Preparation

The dataset to be examined in this Visual Analytics assignment is taken from Singstat on the Resale Flat prices (based on registration date) in Singapore for the period of Y2022.

Upon checking the dataset, the following columns are available to be used in the analysis:

| Variable Name       | Description of Variable                                      |
|------------------------|------------------------------------------------|
| month               | Datetime (Month) "YYYY-MM" of the associated resale price    |
| town                | Area where the resale flat is located in                     |
| flat_type           | Type of flat eg. 3 room or 4 room                            |
| block               | Block number of the resale flat                              |
| street_name         | Street where the resale flat is located                      |
| storey_range        | Range of floors applicable for the resale price              |
| floor_area_sqm      | area in sqm of the flat                                      |
| flat_model          | Indication of whether flat is 'Improved' or 'New Generation' |
| lease_commence_date | The year in which the lease started                          |
| remaining_lease     | Remaining lease in years and months                          |
| resale_price        | Price of the resale flat in SGD                              |

In order to use the data provided more effectively, the following cleaning steps and variable conversions were done using Excel first before application in R to extract and apply the variables in the visualization.

| S/N | Variable | Transformation Step |
|-----|----------|---------------------|
|     |          |                     |
|     |          |                     |
|     |          |                     |

# Approach and Method 

This analytical visualization will attempt to answer the following questions:

1.  How does the resale price vary with the remaining lease years in general?

2.  Does the time of the year affect the resale price of flats in a town?

3.  Does the storey (eg. higher vs lower floors) matter and affect the resale price?

4.  Can a model of factors that influence price be built?

# **Preparing the Visualizations**

```{r}
pacman::p_load(plotly, DT, patchwork, tidyverse, ggstatsplot, readxl, performance, parameters, see, gtsummary, crosstalk, ggdist, gganimate, reshape2, ggridges, ungeviz)

```

```{r}
data <- read_csv("resale-flat-prices/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards-cleaned.csv")

# Filter for Y2022 data only 
resale_flat <- filter(data, year == 2022)
```

## **(A) How does the resale price vary with the remaining lease years in general?**

To examine this, we apply the visual for significant test of correlation between the resale flat prices and the remaining lease months to see if there is a relationship. Using R, ggscatterstats can be used to obtain the correlation plot between resale flat prices and remaining lease months directly.

```{r}
ggscatterstats(
  data = resale_flat, 
  x = remaining_lease_months, 
  y = resale_price,
  marginal = FALSE,
  xlab = 'Remaining Lease (Months)',
  ylab = 'Resale Price'
)
```

## **(B) Does the time of the year affect the resale price of flats in a town?**

To answer this question, we first plot multi-ridge plots to determine the resale flat price distribution across town areas.

```{r}
ggplot(resale_flat,
       aes(x = resale_price, y = town, fill =town)) +
  geom_density_ridges(alpha = 0.5) + 
  theme_ridges() + 
  labs("Resale Flat Prices Across Towns over the Year 2022") + 
  theme(legend.position = "none") + 
  labs(x = "Resale Flat Prices in Y2022", y = "Town") 
```

```{r}
ggplot(resale_flat, 
       aes(x =as.factor(month), resale_price, color = town)) + 
  geom_bar(stat = "identity") + 
  labs(x = "Months of the Year 2022", y = "Resale Flat Price")
```

```{r}
#| fig.dim = c(8,6)
ggbetweenstats(
  data = resale_flat,
  x = month, 
  y = resale_price,
  type = "np",
  messages = FALSE
)
```

## **(C) Does the storey (eg. higher vs lower floors) matter and affect the resale price?**

To observe the distribution of the prices, a bar plot diagram is used to check the distribution. For this, we apply ggplot's barplot to have an idea of the

```{r}
#| fig.dim = c(8,6)
ggplot(resale_flat, 
       aes(x =storey_range, y = resale_price, color = storey_range)) + 
  geom_bar(position = "dodge",
           stat = "summary", 
           fun = "mean") + 
  labs(x = "Storey Range", y = "Resale Flat Price")
```

Oneway ANOVA Test was conducted between storey ranges and the resale price using ggbetweenstats.

```{r}
#| fig.dim = c(8,6)
ggbetweenstats(
  data = resale_flat,
  x = storey_range, 
  y = resale_price,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

To provide visibility and visualization of the uncertainty, we can plot the Hypthetical Outcome Plots (HOPs):

```{r}
ggplot(data = resale_flat, 
       (aes(x = factor(storey_range), y = resale_price))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    linewidth = 0.4, color = "#0072B2", alpha = 1/2) +
  geom_hpline(data = sampler(25, group = storey_range), 
              height = 0.6,
              color = "#D55E00")+
  theme_bw() + 
  transition_states(.draw, 1, 3)
```

Results: Oneway ANOVA test results shows that the differences in mean of resale flat prices across storey ranges are significant; hence there is evidence to show that storey level can affect the price of the resale flat in general.

## **(D) Can a model of factors that influence price be built?**

To examine this, we first apply a multiple linear regression model to see if a model can be constructed using the numerical variables in our data set.

```{r}
model <- lm(resale_price ~ floor_area_sqm + month + lease_commence_date + remaining_lease_months, data = resale_flat)

model
```

We check for multi-collinearity as follows and find that both 'lease_commence_date' and 'remaining_lease_months' have high degree of collinearity and we should re-run the model again while dropping one of these variables

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

To complete model diagnostic checks, run check_model on the model to obtain an overall view of diagnostic parameters like Homogeity of Variance and Normality of Residuals:

```{r}
#| fig.dim = c(8,8)
check_model(model)
```

Re-run of Model with 'lease_commence_date' dropped:

```{r}
model_2 <- lm(resale_price ~ floor_area_sqm + month + remaining_lease_months, data = resale_flat)

model_2
summary(model_2)
```

```{r}
ggcoefstats(model_2, output ="plot")
```

p-values in the model are significant and the F-value is also large enough to indicate that the model is stable. Plot to visualize the model parameters using `ggcoefstats` to see where the values lie with respect to the terms.
